### 2014-11-15 Rscripts/ms_willow_network.Rmd
lots of updates, too many to detail
need to do a better job of keeping track of the changes I make.

---
title: "Plant genetic variation structures herbivore-parasitoid networks"
author: "Matthew Barbour"
date: "20 August, 2014"
output: word_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.